---
title: "Dee_Models2"
author: "Stella Dee"
date: "2022-11-19"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown

###Q1: Re-create the conditional boxplot of penguin body mass conditioned on sex and species. Include your boxplot as a figure in your report.

```{r}
dev.off()
require('palmerpenguins')
par(mfrow = c(1,3)) 
labels= c("female\nAdelie","male\nAdelie","female\nChinstrap","male\nChinstrap","female\nGentoo","male\nGentoo")
boxplot(body_mass_g ~ sex:species, data = penguins, names=labels,ylab= "Body Mass(g)", xlab="", las=2)
boxplot(body_mass_g ~ sex, data = penguins, names=c("female","male"),ylab= "Body Mass(g)", xlab="", las=2)
boxplot(body_mass_g ~ species, data = penguins, names=c("Adelie","Chinstrap", "Gentoo"),ylab= "Body Mass(g)", xlab="", las=2)
```


###Q2: Based on the boxplots, do you think male penguins are significantly heavier than female penguins of the same species? Explain your reasoning, and be sure to explain why you think any observed differences are significant or not.

###Q3: Do you think adding sex to a model that already includes species will improve the model fit?

    Make sure you justify your answer based on the boxplots and not results of a statistical test
    

###Q4: Show the R-code you used to build fit_both.
        Hint: You only need one line of code to do this!
    Q5 (2 pts.): What is the base case for the two-way model that includes sex and species?



    Q6 (2 pts.): What are the names of the two coefficients (from the first column of the coefficient table) that you need to calculate the average mass of female Chinstrap penguins?

    Q7 (2 pts.): What is the predicted average mass of female Chinstrap penguins in the interactive model?

    Q8 (2 pts.): What is the observed average mass of female Chinstrap penguins, calculated from the penguins data?
        Hint: You’ll need to do some logical subsetting to get the answer!
        Hint: [Optional] If you want to be fancy, you can check out the aggregate() function for an elegant approach that does not use logical subsetting.


